---
title: "R Notebook"
output: html_notebook
---


```{r}
library(readr)
job_access_gap <- read_csv("~/Documents/CS - ML/job_access_gap.csv")
# View(job_access_gap)

library(readr)
job_access_gap_dict <- read_csv("~/Documents/CS - ML/job_access_gap_dict.csv")
View(job_access_gap_dict)
```

```{r}
colnames(job_access_gap)
```

```{r}
job_access_gap_dict %>% 
  select(definition)
```
```{r}
nrow(job_access_gap)
```


```{r}
job_access_gap %>% 
  select(OverallJobAccess)
```

Why does there exist values outside [0,1]??
```{r}
job_access_gap %>% 
  ggplot(., aes(x=OverallJobAccess)) + 
  geom_histogram(fill="black") +
  labs(title="Overall Job Access",
       sub="Normalized measure of overall access to jobs for low wage job seekers",
        x ="Job Access", y = "Count") +
  theme_bw()
```
```{r}
job_access_gap %>% 
  filter(OverallJobAccess >= 1)
```



```{r}
job_access_gap %>% 
  filter(OverallJobAccess >= 0 & OverallJobAccess <= 1 ) %>% 
  ggplot(., aes(x=OverallJobAccess)) + 
  geom_histogram(binwidth = 0.02, fill="black") +
  labs(title="Overall Job Access",
       subtitle="Normalized measure of overall access to jobs for low wage job seekers",
        x ="Job Access", y = "Count") +
  theme_bw()
```


```{r}
job_access_gap %>% 
  summarize(Min_OverallJobAccess = min(OverallJobAccess, na.rm=TRUE),
             Q1_OverallJobAccess = quantile(OverallJobAccess, .25, na.rm=TRUE),
             Avg_OverallJobAccess = mean(OverallJobAccess, na.rm=TRUE), 
             Q3_OverallJobAccess = quantile(OverallJobAccess, .75, na.rm=TRUE),
             Max_OverallJobAccess = max(OverallJobAccess, na.rm=TRUE)
             )
```

Why are there values above 1?
```{r}
job_access_gap %>% 
  filter(OverallJobAccess >= 10)
```



The mean job access score is around 0.6.
```{r}
job_access_gap %>% 
  filter(OverallJobAccess >= 0 & OverallJobAccess <= 1 ) %>% 
  summarize(Min_OverallJobAccess = min(OverallJobAccess),
             Q1_OverallJobAccess = quantile(OverallJobAccess, .25),
             Avg_OverallJobAccess = mean(OverallJobAccess, na.rm=TRUE), 
             Q3_OverallJobAccess = quantile(OverallJobAccess, .75),
             Max_OverallJobAccess = max(OverallJobAccess)
             )
```

```{r}
job_access_gap %>% 
  select(lowwagelaborforce)
```


```{r}
job_access_gap %>% 
  ggplot(., aes(x=lowwagelaborforce)) + 
  geom_histogram(fill="black") +
  labs(title="Low Wage Labor Force",
       subtitle="Count of low wage labor force",
        x ="Count of low wage labor force in a block", y = "Count") +
  theme_bw()
```
On avg, there are around 260 low wage laborers in a block. However, there are negative values too.
```{r}
job_access_gap %>% 
  summarize(Min_laborforce = min(lowwagelaborforce, na.rm=TRUE),
             Q1_laborforce = quantile(lowwagelaborforce, .25, na.rm=TRUE),
             Avg_laborforce = mean(lowwagelaborforce, na.rm=TRUE), 
             Q3_laborforce = quantile(lowwagelaborforce, .75, na.rm=TRUE),
             Max_laborforce = max(lowwagelaborforce, na.rm=TRUE)
             )

job_access_gap %>% 
  filter(lowwagelaborforce>=0) %>% 
  summarize(Min_laborforce = min(lowwagelaborforce, na.rm=TRUE),
             Q1_laborforce = quantile(lowwagelaborforce, .25, na.rm=TRUE),
             Avg_laborforce = mean(lowwagelaborforce, na.rm=TRUE), 
             Q3_laborforce = quantile(lowwagelaborforce, .75, na.rm=TRUE),
             Max_laborforce = max(lowwagelaborforce, na.rm=TRUE)
             )
```

```{r}
job_access_gap %>%
  ggplot(., aes(x=Access30Transit)) + 
  geom_histogram(binwidth = 0.02, fill="black") +
  labs(title="Access30Transit",
       subtitle="Normalized measure of jobs accessible within 30 minutes via public transit at peak hours",
        x ="Job Access", y = "Count") +
  theme_bw()
```

Why are there values of normalized scores outside 0-1?
```{r}
job_access_gap %>%
  filter(Access30Transit >= 0 & Access30Transit <= 1 ) %>% 
  ggplot(., aes(x=Access30Transit)) + 
  geom_histogram(binwidth = 0.02, fill="black") +
  labs(title="Access30Transit",
       subtitle="Normalized measure of jobs accessible within 30 minutes via public transit at peak hours",
        x ="jobs accessible score (30 min transit)", y = "Count") +
  theme_bw() 
```
There is a long right tail in this distribution; I wonder which cities have really good access.
The mean job access score is around 0.6.
```{r}
job_access_gap %>% 
  filter(Access30Transit >= 0 & Access30Transit <= 1 ) %>% 
  summarize(Min_Access30Transit = min(Access30Transit),
             Q1_Access30Transit = quantile(Access30Transit, .25),
             Avg_Access30Transit = mean(Access30Transit, na.rm=TRUE), 
             Q3_Access30Transit = quantile(Access30Transit, .75),
             Max_Access30Transit = max(Access30Transit)
             )
```

The sample size for Seattle is the biggest.
```{r}
job_access_gap %>% 
  filter(spatialmismatch > 0 & spatialmismatch <= 1 ) %>% 
  ggplot(., aes(x=spatialmismatch))+
  geom_histogram(binwidth = 0.01, color="red")+
  theme_bw() +
  facet_wrap(~MSA)
```

