---
title: "2019-person-EDA"
author: "Albert Sun"
date: "2/17/2022"
output: html_document
---

```{r include=FALSE}
library(tidyverse)
library(readr)
```

```{r include=FALSE}
df_raw <- read_csv('../../../data/17-19survey-person.csv')
glimpse(df_raw)
```

```{r}
household <- read_csv('../../../data/2017-2019-pr2-1-Household.csv')
glimpse(household)
```

```{r}
household%>%
  select(prev_home_rgcname) 
```

```{r}
household%>%
  select(final_home_rgcnum) 

```

```{r}
# queries for suburbanised people

stay_in_washington <- household %>%
  filter(sample_county == 'King') %>%
  filter(prev_home_wa == 'Yes, previous home was in Washington')

stay_in_washington
```

Here are data columns that are particularly interesting: 
- `final_home_tract` - census tract of where the final home is located
- `prev_home_tract` - census tract of where the first home is located

The breakdown for FIPS is 
- 2 - State
- 3 - County
- 6 - Tract
- 1 - Block Group

This website allows you to see the boundaries of specific census tracts: 

https://www.policymap.com/newmaps#/

This leads to the question - how do I categorize tracts into larger groups? 

- GIS of all census tracts in King County: https://www5.kingcounty.gov/sdc/Metadata.aspx?Layer=tracts10_shore


```{r}
household %>%
  select(prev_home_wa, hhid, final_home_tract, prev_home_tract, final_home_rgcnum, final_home_uvnum, prev_home_rgcname)
```

It seems as if the census tracts are only listed if the person lived in washington before; i.e. prev_home_wa == "Yes, previous home was in Washington". Thus, I think it is ok to filter out for answering yes as I did in the code chunk before this one. 

```{r}
stay_in_washington %>%
  select(hhid, final_home_tract, prev_home_tract, final_home_rgcnum, final_home_uvnum, prev_home_rgcname)
```

```{r}
df_prev_home <- stay_in_washington %>%
  group_by(prev_home_tract) %>%
  summarise(n = n()) 

df_prev_home
```

Top locations are: 
- Redmond, 
- Seattle West Edge/Central Business District, 
- Seattle Bell Town, 
- Capitol Hill, 
- Between University District/Wallingford

```{r}
df_final_home <- stay_in_washington %>%
  group_by(final_home_tract) %>%
  summarise(n = n())

df_final_home
```

Top locations are: 
- Redmond, 
- Northgate (Northern Seattle)
- Between University District/Wallingford
- Southlake Union (near airport)
- Capitol Hill

In order to track suburbanisation in this dataset, I think we would need a systematic way to categorize whether a specific census tract is suburban or not. Having a dataset that has a list of all the census tracts and what type of locations they describe would be a good start to doing this. 

```{r joining the tracts}
df_home_tracts_2019 <- full_join(df_prev_home, df_final_home, 
                         by=c('prev_home_tract' = 'final_home_tract')) %>%
  rename(tract = prev_home_tract,
         num_prev_2019 = n.x,
         num_final_2019 = n.y) %>%
  mutate(tract = as.character(tract))

df_home_tracts_2019

write_csv(df_home_tracts_2019, "../../../data/survey-household-tracts-moving.csv")
```

We just did that for the 2019 survey. Now, let's do it for the 2017 survey. 

```{r do the same t}
library(readxl)
library(janitor)

df_stay_in_washington_2017 <- read_excel('../../../data/2017-pr2-1-household.xlsx') %>%
  row_to_names(row_number = 1) %>%
  select(prev_home_wa, prev_home_tract, final_home_tract) %>%
  filter(prev_home_wa == 1)
```

```{r}
df_stay_in_washington_2017

df_prev_home_2017 <- df_stay_in_washington_2017 %>%
  group_by(prev_home_tract) %>%
  summarise(n = n()) 

df_final_home_2017 <- df_stay_in_washington_2017 %>%
  group_by(final_home_tract) %>%
  summarise(n = n())

df_home_tracts_2017 <- full_join(df_prev_home_2017, df_final_home_2017, 
                         by=c('prev_home_tract' = 'final_home_tract')) %>%
  rename(tract = prev_home_tract,
         num_prev_2017 = n.x,
         num_final_2017 = n.y)

df_home_tracts_2017
```

```{r join 2017 and 2019 tracts and output csv}
df_home_tracts_total <- full_join(df_home_tracts_2017, df_home_tracts_2019, 
                         by='tract') 
df_home_tracts_total

write_csv(df_home_tracts_total, "../../../data/survey-household-tracts-moving.csv")
```
